Comparison of the solution of the shallow equation with experimental

### 2015-03-10 doc/paper/paper.tex
update paper.tex

\begin{abstract}
In this paper we compare the solution of the shallow water equations with experimental results. The experiment simulates the interaction between an incident bore and a free-standing coastal structure. The shallow water equations are solved using the CLAWPACK software.
\end{abstract}

@@

\section{Dam Break with square column}\label{Sec:Square}
Figure \ref{fig:waveheight_square_x=11.0} gives a comparison of the time history of the wave height measured at 0.04m ahead of the leading edge of the square column.(0.1m ahead of center of the column)
Note that this figure is from t=3.0 to t=4.0, which is quite a short range compared to the previous figures.
There are two jumps in result given by geoclaw. First jump change water level from around 0.02 to 0.1, second one jump from 0.1 to a higher level of around 0.22.
If we recall that the wave height is around 0.1 in \ref{fig:waveheight_nocolumn}, we can confirm this first jump is cause by the arrival of shock wave caused by dambreak. The second jump is induced by wave reflected by the column.  
If we compare the time interval between first jump and second jump with that in \ref{fig:waveheight_square_x=11.02}, we can see that as the gauge gets closer to column, this time interval also gets shorter.
The arrival time of the peak predicted by geoclaw is about 0.2 second earlier than the time measured from experiment and in experiment, the jump is more like a a smooth increase. 
This can be explained by the difference between the shallow water equations that geoclaw solves and N-S equations that better describe the physics.
In shallow water equations, fluid viscosity is not taken into account, while it's significant in N-S equation (or real physics).
This causes the shock remain its shape in as it propagates in shallow water equation and smooth out in experiment.
Thus we see a jump in result from shallow water equation while a smooth increase in experiment result.
Shallow water equation is essentially a hyperbolic PDE, whose solution depends on initial condition along characteristics in spatio-temporal domain. 
Thus before the propagating shock arrive, all water downstream the shock stays still. In other words, it even does not know there is a column before the shock hits the column. In contrast, N-S equation is a elliptic PDE. 
In real physics govened by N-S equation, when front of the shock wave gets close to the column, water downstream it and upstream column are affected and resist the comming of wave to some extent, which raises the wave hegiht even before the wave hits the column.   
This explains why we cannot see two jumps in experiment results.
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x11}
    \caption{time history of wave height for case with square column}
    \label{fig:waveheight_square_x=11.0}
\end{figure}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x1102}
    \caption{time history of wave height for case with square column, at x = 11.02}
    \label{fig:waveheight_square_x=11.02}
\end{figure}

\section{Dam Break with cylinder column}\label{Sec:Cylinder}
Figure \ref{fig:waveheight_cylinder_x=11.0} is the time history of the wave height measured at 0.04m ahead of the leading edge of cylinder column.(0.1m ahead of the center of the column)
Since the way geoclaw treat a cell partly belong to cylinder and partly belong to tank bottom is to give that cell an average value of topography from cylinder and tank bottom based on area ratio, we get a quasi-circle actually instead of circle. Thus the result in this case does not agree well with the experiment result. 
Future test can be conducted by setting a higher level for mesh refinement to get a better circle.
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.5\textwidth]{./plots/waveheight_cylinder_x11}
    \caption{time history of wave height for case with cylinder column}
    \label{fig:waveheight_cylinder_x=11.0}
\end{figure}

>>>>

\begin{abstract}
In this paper we compare the solution of the shallow water equations with experimental results. The experiment simulates the interaction between an incident bore and a free-standing coastal structure. The shallow water equations are solved using the CLAWPACK software.
\par
Three different cases are simulated and compared. In the first case, the column is removed. 
This is a simple dambreak problem and history of wave height and velocity at where column center should be are measured and compared. 
In the second case, a square column is set and wave height at some locations ahead the column are measured and compared. 
For the third case, the square column is replaced with a cylinder with the same data measured and compared. 
The discrepancy between geoclaw simulation and experiment results are also discussed and analyzed in a deep manner.
\end{abstract}

@@

\section{Dam Break with square column}\label{Sec:Square}
Figure \ref{fig:waveheight_square_x=11.0} gives a comparison of the time history of the wave height measured at 0.04m ahead of the leading edge of the square column.(0.1m ahead of center of the column)
Note that this figure is from t=3.0 to t=4.0, which is quite a short range compared to the previous figures.
There are two jumps in result given by geoclaw. First jump change water level from around 0.02 to 0.1, second one jump from 0.1 to a higher level of around 0.22.
If we recall that the wave height is around 0.1 in figure \ref{fig:waveheight_nocolumn}, we can confirm this first jump is cause by the arrival of shock wave caused by dambreak. The second jump is induced by wave reflected by the column.  
If we compare the time interval between first jump and second jump with that in figure \ref{fig:waveheight_square_x=11.02}, we can see that as the gauge gets closer to column, this time interval also gets shorter.
There are two distinct differences between geoclaw result and experiment result. First, in experiment, the jump is more like a smooth increase. Second, arrival time of the peak predicted by geoclaw is about 0.08 second earlier than the time measured from experiment.  
The difference in shock shape can be explained by the difference between the shallow water equations that geoclaw solves and N-S equations that better describe the physics.
In shallow water equations, fluid viscosity is not taken into account, while it's significant in  real physics.
This causes the shock remain its shape as it propagates in shallow water equation while smooth out in experiment.
Thus we see a jump in result from shallow water equation while a smooth increase in experiment result.
If we zoom in figure \ref{fig:waveheight_nocolumn} for smaller time range from 3s to 4s, we get figure \ref{fig:waveheight_nocolumn_zoomin}.
In figure \ref{fig:waveheight_nocolumn_zoomin}, arrival time of peak predicted by geoclaw is almost the same as that from experiment (only about 0.02 second earlier).  
This time interval is much smaller than that in figure \ref{fig:waveheight_square_x=11.0} though data are not measured at exactly the same point for this two plots. 
This difference indicates that the column gives more resistance to shock waves in experiment than it does in shallow water equation model. 
A simple analysis to this phenomenon is draw as below.
Shallow water equation is essentially a hyperbolic PDE, whose solution depends on initial condition along characteristics in spatio-temporal domain. 
Thus before the propagating shock arrive, all water downstream the shock stays still. 
In other words, it even does not know there is a column before the shock hits the column.
This can be verified by checking history of wave height measured at exactly the same place, 0.04m ahead of leading edge of the square column, but in different cases, one with a square column and one without.
This comparison is showed in figure \ref{fig:waveheight_x11_comparison}. As we expected, there is no difference in arrival time of the shock wave.
In contrast, N-S equation is a elliptic PDE. 
In real physics governed by N-S equation, when front of the shock wave gets close to the column, water downstream it and upstream column are affected and resist the comming of wave to some extent. 
This slows down the propagating speed of shock wave when it comes close to the column in real world and thus shifts the peak toward later time in the plots.
    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x11}
    \caption{time history of wave height for case with square column}
    \label{fig:waveheight_square_x=11.0}
    \includegraphics[width=0.5\textwidth]{./plots/waveheight_nocolumn_zoomin}
    \caption{time history of wave height for case without column (zoom in)}
    \label{fig:waveheight_nocolumn_zoomin}
    \includegraphics[width=0.5\textwidth]{./plots/waveheight_x11_comparison}
    \caption{comparison of time history of wave height for case with a square column and case without one}
    \label{fig:waveheight_x11_comparison}
    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x1102}
    \caption{time history of wave height for case with square column, at x = 11.02}
    \label{fig:waveheight_square_x=11.02}
Figure \ref{fig:waveheight_square_x=11.04} shows time history of wave height measured at front face of the square column. 
Difference between arrival time of peak still exists as expected. After the peak arrive, the wave gradually becomes stable as indicated by a horizontal line in the plots, in which region the differences between geoclaw and experiment are quite small. 
    \includegraphics[width=0.5\textwidth]{./plots/waveheight_square_x1104_largerTimeRange}
    \caption{time history of wave height for case with square column, at x = 11.04}
    \label{fig:waveheight_square_x=11.04}

\section{Dam Break with cylinder column}\label{Sec:Cylinder}
Figure \ref{fig:waveheight_cylinder_x=11.0} is time history of the wave height measured at 0.04m ahead of the leading edge of cylinder column.(0.1m ahead of the center of the column)
%Since the way geoclaw treat a cell partly belong to cylinder and partly belong to tank bottom is to give that cell an average value of topography from cylinder and tank bottom based on area ratio, we get a quasi-circle actually instead of circle. Thus the result in this case does not agree well with the experiment result. 
%Future test can be conducted by setting a higher level for mesh refinement to get a better circle.
Similar feature as what we see in case with a square column can be found in the figure.
However, the wave height at second peak in this case is almost the same as that in square-column case (around 0.22m in both case), while this feature is quite different for experiment results between two cases.
\par
Note that diameter of the cylinder is equal to length of the transection for square column. Thus the square column can block the flow in in a more efficient way than cylinder does since cylinder has a smoother shape for the fluid to pass by. 
This explains why we see a lower peak of wave height as well as less steep curve for history of wave height in case with cylinder in experiment. 
In geoclaw, a geometry is represented with many quadrangular cells. In our cylinder case, the cylinder is represented with many square cells since our background mesh is simple structured quadragular mesh with square cells. 
Based on this, we can deduce in geoclaw, when the shock wave hits the cylinder, it is hitting onto many small faces vertical to flow direction at different locations. 
Thus it is similar to hitting onto one big face vertical to flow direction, which is the case with square column.  
This is why my get similar result in geoclaw for case with square column and case with cylinder.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.5\textwidth]{./plots/waveheight_cylinder_x11}
    \caption{time history of wave height for case with cylinder column}
    \label{fig:waveheight_cylinder_x=11.0}
\end{figure}
\par
Figure \ref{fig:waveheight_cylinder_x=11.04} is time history of wave height for cylinder case. As we expected, peak predicted by geoclaw appears earlier but the peak values agree quite well.
After the peak, geoclaw gives a satisfied prediction of wave height.
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.5\textwidth]{./plots/waveheight_cylinder_x1104_largerTimeScale}
    \caption{time history of wave height for case with cylinder column, at x = 11.04}
    \label{fig:waveheight_cylinder_x=11.04}
\end{figure}
